### 2016-05-17 text/new_formal_background.tex
Work on defining model transformation.

For simplification purposes, we will not model the edge cardinalities given by a metamodel in our notion of typed graph. Note that in fact we require these cardinalities to be relaxed to perform our graph rewriting.

@@

@@

\bentley{Do we need to have V and E for all vertices and edges in the IOM?}
Additionally we have that $V=V'\cup V''$, $E\subseteq E'\cup E''$ and $\tau\subseteq \tau'\cup \tau''$, where the co-domain of $\tau$ is the union of the co-domains of $\tau'$ and $\tau''$ and the set $\{trace\}$.

@@

\subsection{Semantic Transformation Constructs}
\label{sec:DSLTrans_semantics}

In the definition that follows we introduce the notion of execution of a DSLTrans model transformation. For our purposes it is sufficient to introduce it as an input-output model (see \cref{def:input_output_model}), containing the input model for the transformation, the produced output, and the traceability links built during execution. Due to space limitations, we cannot introduce the semantics of DSLTrans in the main text of this paper. We thus refer the reader to \cref{sec:DSLTrans_formal} for a formal description on how DSLTrans transformation executions are built.

\reviewer{Definition 17: There is a copy/paste error w.r.t. Def. B.16 (which only serves to
strengthen my point, made above, about avoiding this kind of duplication). In
this case, B.16 is the correct version.}

\begin{definition} {Model Transformation Execution\\}
\label{def:modeltransformation}

Let  $tr\in \textsc{Transf}^{sr}_{tg}$ be a transformation and $input \in \textsc{Model}^{sr}$ be a model. Assume we also have that: $$\langle V,E,st,\tau,input,\epsilon\rangle,tr \xrightarrow{trstep} \langle V',E',st',\tau',input,output\rangle$$
A model transformation execution is the input-output model $\langle V',E',st',\tau',input,output\rangle\in \textsc{Iom}^{sr}_{tg}$, where $output\in \textsc{Iom}^{sr}_{tg}$ is an input-output model. The set of all model transformation executions for transformation $tr$ is written $\textsc{Exec}(tr)$. A model transformation with an empty input model is noted $\epsilon_{ex}$. Note that relation $trstep$ is formally defined in \cref{sec:DSLTrans_formal_appendix}.
\end{definition}

Note that ${VT}_{Input}$ and ${ET}_{Input}$ will

Finally, as stated in \cref{def:modeltransformation}, we consider a model transformation execution to be the input-output model (IOM) resulting from executing a set of rules on a starting IOM. This starting IOM includes the transformation's input in its input part and has an empty output part. The starting IOM represents the first step of the transformation when no rule has been executed yet. A transformation execution results from executing all the rules in a DSLTrans model transformation.

\subsubsection{Transformation Language Semantics}

We will now address the semantics of the DSLTrans language. We will start by defining a match function that, given an input-output model and a transformation rule, returns all subgraphs of that input-output model where the rule's match pattern (including backward links) is found.

@@

@@

\begin{definition} {Model Transformation Execution}
\label{def:modeltransformation_appendix}

\CatchFileBetweenTags{\modeltransformation}{text/definitions}{modeltransformationappendix}{\modeltransformation}
\end{definition}

\CatchFileBetweenTags{\modeltransformationtextappendix}{text/definitions}{modeltransformationtextappendix}{\modeltransformationtextappendix}

>>>>

For simplification purposes, we will not represent edge cardinalities or containment relationships given by a metamodel in our notion of typed graph. Note that in fact we require these conditions to be relaxed to perform our graph rewriting.

@@

\bentley{I believe that the indirect edges don't need to be removed. There's a bit of handwaving, so maybe there are edge duplicates?}

@@

Let $\mathit{IOM}$ be the set of all input-output models.

Note that the $\VT$ and $\ET$ for input and output will come from the input and output metamodels. \bentley{Expand}

\bentley{Do we need to have V and E for all vertices and edges in the IOM? I don't think we talk about all vertices or edges in an IOM.}
Additionally we have that $V=V'\cup V''$, $E\subseteq E'\cup E''$ and $\tau\subseteq \tau'\cup \tau''$, where the co-domain of $\tau$ is the union of the co-domains of $\tau'$ and $\tau''$ and the set $\{trace\}$.

@@

\subsection{Transformation Semantics}
\label{sec:DSLTrans_semantics}

This section will discuss the semantics of a DSLTrans transformation. Given an input model and a transformation, an output layer will be produced through the repeated execution of rules and layers.

\subsubsection*{Execution of a Model Transformation}

We will begin our discussion of the semantics of model transformation by discussing the conditions for executing a model transformation on an input model.

\begin{definition} {Execution of a Model Transformation\\}
\label{def:modeltransformation}

The execution of a model transformation is a function $transform: IOM \times Transforms \rightarrow IOM$.

For a well-formed transformation execution, we require the IOM in the domain to contain only an input graph. Recall that a IOM is the three-tuple $\big\langle \mathit{Input}$, $\mathit{Output}$, $\mathit{trace}\big\rangle$. We therefore require that both the $\mathit{Output}$ and $\mathit{trace}$ components in the domain IOM to be empty. This IOM thus represents the first step of the transformation, where no rule has been executed yet.

The IOM in the co-domain will contain this input graph, as well as the output graph and traceability links produced by the execution of the transformation. Let $\textsc{Execs}_t$ be the set of all well-formed IOMs produced for transformation $t$.

\subsubsection*{Execution of a Rule}

We will now address the execution of a rule to in the DSLTrans language.

We will start by defining a match function that, given an input-output model and a transformation rule, returns all subgraphs of that input-output model where the rule's match pattern (including backward links) is found.

@@

\subsubsection*{Layer Step Semantics}

@@
